Learning action models with minimal observability

### 2019-05-02 doc/aij18/minor-revision/aij18.tex
AIJ18-minor: rephrasing some points
AIJ18-minor: final version

\title{Learning action models with minimal observability}
\author[label1]{Diego Aineto}
\author[label1]{Sergio Jim\'{e}nez Celorrio}
\author[label1]{Eva Onaindia}
\address[label1]{Department of Computer Systems and Computation, Universitat Polit\`ecnica de Val\`encia. Spain}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{<author name>}
%% \address[label1]{<address>}
%% \address[label2]{<address>}

@@

In this paper we have presented \FAMA, an approach for learning \strips\ action models based on the compilation of the learning task into a planning task. The distinctive characteristic of \FAMA over other state-of-the-art approaches is its ability to learn from minimal input knowledge and, more particularly, from minimal observability plan traces with no observed actions and no observed intermediate states. By relaxing the input constraints, \FAMA opens up the way for action model learning to operate on real-world problems, as opposite to current approaches where the heavy input restrictions have limited their applicability to synthetic benchmarks. Our approach is thus very well suited for learning action models in data-based applications where the only observable information is a possibly incomplete sequence of partially observed states.

@@

%This great flexibility of \FAMA to different amounts of minimal input knowledge impacts the complexity of the planning tasks. When the length of the input plan traces is unknown, that is, we ignore the number of steps of the decision-making process of the agent, \textcolor{red}{the learning task to be solved can no longer be addressed using current approaches}. This is one of the reasons that justifies our compilation-to-planning scheme in contrast to the most extended SAT-solving scheme used in many existing learning systems.

>>>>

\title{Learning action models with minimal observability}
\author[label1]{Diego Aineto\corref{cor1}}
\ead{dieaigar@dsic.upv.es}
\author[label1]{Sergio Jim\'{e}nez Celorrio}
\ead{serjice@dsic.upv.es}
\author[label1]{Eva Onaindia}
\ead{onaindia@dsic.upv.es}

\cortext[cor]{Corresponding author}

\address[label1]{Department of Computer Systems and Computation, Universitat Polit\`ecnica de Val\`encia. Camino de Vera s/n. 46022 Valencia, Spain}

@@

In this paper we have presented \FAMA, an approach for learning \strips\ action models based on the compilation of the learning task into a planning task. The distinctive characteristic of \FAMA over other state-of-the-art approaches is its ability to learn from minimal observability, i.e., plan traces with no observed actions and no observed intermediate states. By relaxing the input constraints, \FAMA opens up the way for action model learning to operate on real-world problems, as opposite to current approaches where the heavy input restrictions have limited their applicability to synthetic benchmarks. Our approach is thus suitable for learning action models in data-based applications where the only observable information is a possibly incomplete sequence of partially observed states.

@@

### 2019-05-02 doc/aij18/minor-revision/background.tex
AIJ18-minor: final version

In the following, we present a comprehensive insight of the particularities of the seven systems presented in Table \ref{table:models_comparison1} and Table \ref{table:models_comparison2}. This exposition will also help us to highlight in section \ref{task_definition} the value of our contribution \FAMA.

>>>>

In the following, we present a comprehensive insight of the particularities of the eight systems presented in Table \ref{table:models_comparison1} and Table \ref{table:models_comparison2}. This exposition will also help us to highlight in section \ref{task_definition} the value of our contribution \FAMA.

### 2019-05-02 doc/aij18/minor-revision/evaluation.tex
AIJ18-minor: final version

We can observe in Table \ref{table:models_comparison2} that most of the approaches use a similar syntax-based metric that consists in (1) counting the missing and extra fluents that appear in the learned model wrt the GTM and (2) normalizing this error by the total number of all the possible preconditions and effects of an action model. This is an \emph{optimistic} metric since error rates are not normalized by the size of the actual GTM. The set of preconditions and effects of the GTM is usually smaller than the set of all possible preconditions and effects and thereby it turns out that these syntax-based metrics may output error rates below 100\% for totally wrong learned models. To overcome this limitation we propose to use two standard metrics from ML, {\em precision} and {\em recall}, that are frequently used in pattern recognition, information retrieval and binary classification~\cite{davis2006relationship}. These two syntactic metrics are generally more informative than counting the number of errors between the learned action models and the GTM:

>>>>

\textcolor[rgb]{1.00,0.00,0.00}{As commented before, the large majority of approaches shown in Table \ref{table:models_comparison2} evaluate the learned action models wrt a GTM. These GTM are also  used to generate the learning plan traces and they are normally domain definitions taken from the IPC.} We can also observe in Table \ref{table:models_comparison2} that most of the approaches use a similar syntax-based metric that consists in (1) counting the missing and extra fluents that appear in the learned model wrt the GTM and (2) normalizing this error by the total number of all the possible preconditions and effects of an action model. This is an \emph{optimistic} metric since error rates are not normalized by the size of the actual GTM. The set of preconditions and effects of the GTM is usually smaller than the set of all possible preconditions and effects and thereby it turns out that these syntax-based metrics may output error rates below 100\% for totally wrong learned models.

In order to overcome this limitation we propose to use two standard metrics from ML, {\em precision} and {\em recall}, that are frequently used in pattern recognition, information retrieval and binary classification~\cite{davis2006relationship}. These two syntactic metrics are generally more informative than counting the number of errors between the learned action models and the GTM:

### 2019-05-02 doc/aij18/minor-revision/introduction.tex
AIJ18-minor: final version

FAMA is a new learning approach, based on AI planning technology, that automatically compiles the task of learning \strips\ actions into a planning task which is then solved with a planner~\cite{aineto2018learning}. The construction of a \strips\ planning model starts out from a set of plan traces containing the observation of several plan executions. As mentioned above, a plan trace may comprise none of the actions executed by the agent but must include, at least, the initial state $s_0$ and final state $s_f$ of the execution. The compilation scheme lies in defining a planning task from the set of plan traces using a set of \emph{building actions} that insert the preconditions and effects of a learned action, and a set of \emph{validating actions} that validate the learned actions in the plan traces. Hence, a solution to this planning task is a plan that determines the preconditions and effects of the actions of a \strips\ planning model $M$ while ensuring consistency with the input traces. We say that a model $M$ is \emph{consistent} with a plan trace when $M$ can produce a solution $\pi$ to the planning problem $\tup{s_0,s_f}$ so that: (1) $\pi$ contains the observed actions of the plan trace, if any, and (2) the states generated by the application of $\pi$ to $s_0$ will encompass all the (possibly) partially observed states of the trace.

>>>>

FAMA is a new learning approach, based on AI planning technology, that automatically compiles the task of learning \strips\ actions into a planning task which is then solved with a planner~\cite{aineto2018learning}. The construction of a \strips\ planning model starts out from a set of plan traces containing the observation of several plan executions. As mentioned above, a plan trace may comprise none of the actions executed by the agent but must include, at least, the initial state $s_0$ and final state $s_n$ of the execution. The compilation scheme lies in defining a planning task from the set of plan traces using a set of \emph{building actions} that insert the preconditions and effects of a learned action, and a set of \emph{validating actions} that validate the learned actions in the plan traces. Hence, a solution to this planning task is a plan that determines the preconditions and effects of the actions of a \strips\ planning model $M$ while ensuring consistency with the input traces. We say that a model $M$ is \emph{consistent} with a plan trace when $M$ can produce a solution $\pi$ to the planning problem $\tup{s_0,s_n}$ so that: (1) $\pi$ contains the observed actions of the plan trace, if any, and (2) the states generated by the application of $\pi$ to $s_0$ will encompass all the (possibly) partially observed states of the trace.

### 2019-05-02 doc/aij18/minor-revision/learning_task_def.tex
AIJ18-minor: minor fixes
AIJ18-minor: final version

\begin{footnotesize}
	\begin{verbatim}
	;;; State observation
	(clear A) (on A B) (ontable B)
	\end{verbatim}
\end{footnotesize}

@@

\item \textbf{$\tau$ does not identify the length of $\pi$}. This happens when $\tau_a$ and $\tau_s$ are both partially observed (\PO) or non-observable (\NO) action/state trajectories. In this case, we are unaware of the number of actions of $\pi$ and the number of states induced by $\pi$. This gives rise to a completely different scenario and a more challenging learning task that brings one key difference: the transition between two given observed states of $\tau$ may now involve more than one action; i.e., $\theta(s_i,\tup{a_1,\ldots,a_k})=s_{i+1}$, with $k \geq 1$, $k$ unknown and unbounded, and so \textbf{the horizon of $\pi$ is no longer known}. This justifies the use of \textbf{planning techniques} for solving the learning task, which can now be interpreted as \emph{filling the gap} between two observable points of $\tau$. \textcolor{red}{SAT approaches, on the other hand, are no longer obviously applicable given the lack of a length bound}. In this particular scenario, the actual number of plans consistent with the given observed plan trace is also unbounded and grows exponentially with the actual length of the plans (that is now unknown). Therefore, when we assume partial observability in both actions and states, a learning approach must consider that the length of the observed plan traces is not an indication of the actual length of the plan, which motivates and justifies the use of planning, as our proposal of compiling the learning task to a classical planning problem.

>>>>

\begin{footnotesize}
	\begin{verbatim}
	;;; Final state observation
	(clear A) (on A B) (ontable B)
	\end{verbatim}
\end{footnotesize}

@@

\item \textbf{$\tau$ does not identify the length of $\pi$}. This happens when $\tau_a$ and $\tau_s$ are both partially observed (\PO) or non-observable (\NO) action/state trajectories. In this case, we are unaware of the number of actions of $\pi$ and the number of states induced by $\pi$. This gives rise to a completely different scenario and a more challenging learning task that brings one key difference: the transition between two given observed states of $\tau$ may now involve more than one action; i.e., $\theta(s_i,\tup{a_1,\ldots,a_k})=s_{i+1}$, with $k \geq 1$, $k$ unknown and unbounded, and so \textbf{the horizon of $\pi$ is no longer known}. This justifies the use of \textbf{planning techniques} for solving the learning task, which can now be interpreted as \emph{filling the gap} between two observable points of $\tau$. \textcolor{red}{SAT approaches, on the other hand, are no longer straightforwardly applicable given the lack of a length bound for the observed plan trace}. In this particular scenario, the actual number of plans consistent with the given observed plan trace is also unbounded and grows exponentially with the actual length of the plans (that is now unknown). Therefore, when we assume partial observability in both actions and states, a learning approach must consider that the length of the observed plan traces is not an indication of the actual length of the plan, which motivates and justifies the use of planning, as our proposal of compiling the learning task to a classical planning problem.
